---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Importing beef sales by pounds into file. 

```{r}
library(tidyverse)
library(forecast)
library(lubridate)
library(readxl)
USDA <- read_excel('D:/Capstone/Timeseries Analysis/Beefsalesbyregionusda.xlsx')
```
Consolidating by month for total national beef 

```{r}
beef <- USDA%>%
  rename('Date'= 'Report Date',
         'Nationalpounds'='National in Pounds',
         'Westpounds'='Westcoast in Pounds')%>%
  arrange(Date)%>%
  select(Date, Month,Westpounds )%>%
  mutate(Month=format(Month, "%m-%Y"))%>%
  group_by(Month)%>%
  summarize(National=sum(Westpounds))

beef

```

I had to export it to Excel so I can organize the dates properly
```{r}
write.csv(beef, 'D:/Capstone/Timeseries Analysis/westbeef.csv', row.names = TRUE)
```

Reading my organized file back into R
```{r}
beef <- read_excel('D:/Capstone/Timeseries Analysis/westbeef.xlsx')
```

```{r}
beef.ts.month <- ts(beef$National, start = c(2021,9), end = c(2024,8), frequency = 12)
plot(beef.ts.month)
```


```{r}
nvalid <- 6
ntrain <- length(beef.ts.month) - nvalid
train.ts <- window(beef.ts.month, start = c(2021,9), end = c(2021,ntrain))
valid.ts <- window(beef.ts.month, start= c(2021, ntrain+1), end = c(2021,ntrain+nvalid))
```

```{r}
plot(train.ts)
```
```{r}
Acf(beef.ts.month)
```

There a couple points of white noise. But lets see how the Models look.

Plot shows seasonality and some linear trend. Build model that captures both.


```{r}
reg_linear_season <- tslm(train.ts ~ trend + season)

reg_linear_season_pred <- forecast(reg_linear_season, h=nvalid, level = 0)

plot(reg_linear_season_pred, include = 20)
lines(valid.ts, lwd=2)
res_valid <- valid.ts - reg_linear_season_pred$mean # forecast error = true value - forecasted value

data.frame(actual = valid.ts, forecasted = reg_linear_season_pred$mean, forecast_error=res_valid)
accuracy(reg_linear_season_pred$mean, valid.ts)

```
```{r}
poly_season_mod <- tslm(train.ts ~ trend + I(trend^2) + season)
options(scipen = 999)
summary(poly_season_mod)
poly_season_mod_pred <- forecast(poly_season_mod, h=nvalid, level = 0)
head(poly_season_mod_pred$mean)

```

simple exponential smoothing

```{r}
ses_pred <- ses(train.ts, h=nvalid)
head(ses_pred$mean)
```
Next, we also consider a Holt-Winter method

Holt Winter method

```{r}
hw_pred <- hw(train.ts, h=nvalid)
head(hw_pred$mean)
```

```{r}
ets_aan <- ets(train.ts, model = "AAN")
ets_aan_pred <- forecast(ets_aan, h=nvalid, level = 0)
head(ets_aan_pred$mean)
head(ses_pred$mean)
```
ETS Model

```{r}
ets_r <- ets(train.ts)
summary(ets_r)
ets_r_pred <- forecast(ets_r, h=nvalid, level = 0)
head(ets_r_pred$mean)
```
Naive Method

```{r}
snaive_pred <- snaive(train.ts, h=nvalid, level = 0)
```


```{r}
plot(poly_season_mod_pred, include = 24)
lines(ets_aan_pred$mean, col='red', lwd=2)
lines(hw_pred$mean, col='orange', lwd=2)
lines(ets_r_pred$mean, col='pink', lwd=2)
lines(snaive_pred$mean, col='green', lwd=2)
lines(reg_linear_season_pred$mean, col='blue', lwd=2)
lines(valid.ts,col='purple', lwd=2)
lines(poly_season_mod_pred$mean, col='black',lwd=2)
```
```{r}
accuracy(ets_aan_pred)
accuracy(ets_r_pred)
accuracy(poly_season_mod_pred)
accuracy(reg_linear_season_pred)
accuracy(ses_pred)
accuracy(snaive_pred)
```
The best model is Poly season Mod

Now that we select ETS(M,A,A) as our model, we'll apply this to the entire data set ridership.ts to compute the 'future'forecast

```{r}
reg_linear_season_pred_all <- tslm(train.ts ~ trend + season)
nFuture <- 60
reg_linear_season_forecast <- forecast(reg_linear_season_pred_all, h=nFuture, level=95)
reg_linear_season_forecast
plot(reg_linear_season_forecast, include = 24)
```

```{r}
res_valid <- valid.ts - reg_linear_season_pred$mean # forecast error = true value - forecasted value

data.frame(actual = valid.ts, forecasted = reg_linear_season_pred$mean, forecast_error=res_valid/valid.ts)
```

```{r}
print(reg_linear_season_forecast)
```
```{r}

accuracy(reg_linear_season_pred)
accuracy(snaive_pred)
```
```{r}
reg_linear_season_pred <- forecast(reg_linear_season, h=nvalid, level = 0)

reg_linear_season_pred
plot(reg_linear_season_pred, include = 24)
lines(valid.ts,col='green', lwd=2)
```
```{r}
print(reg_linear_season_forecast)

write.csv(reg_linear_season_forecast, 'D:/Capstone/Timeseries Analysis/reglinforecast.csv', row.names = TRUE)

```

